Analysis of Stocks
===================
By Alok Deo Singh


Abstract
========


Initial Preparation
-------------------
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
options(scipen = 999)

library(ggplot2)
library(dplyr)
library(tidyr)
library(GGally)
library(TTR)

```

Loading the Data
----------------
#  Declaring Global Variables
```{r}
companies <- list('axis','hcl','hdfc','icici','indus','infy','kotak','sbi',
                  'tcs','techm','wipro','yes')
file_loc <- list(axis = "Data_Set/AXISBANK.NS.csv",
                 hcl =  "Data_Set/HCLTECH.NS.csv",
                 hdfc = "Data_Set/HDFCBANK.NS.csv",
                 icici = "Data_Set/ICICIBANK.NS.csv",
                 indus = "Data_Set/INDUSINDBK.NS.csv",
                 infy = "Data_Set/INFY.NS.csv",
                 kotak = "Data_Set/KOTAKBANK.NS.csv",
                 sbi = "Data_Set/SBIN.NS.csv",
                 tcs = "Data_Set/TCS.NS.csv",
                 techm = "Data_Set/TECHM.NS.csv",
                 wipro = "Data_Set/WIPRO.NS.csv",
                 yes = "Data_Set/YESBANK.NS.csv")
stock <- NULL
```

#  Load Data frames
```{r echo=FALSE, message=TRUE, warning=TRUE, Load_Data}

load_data <- function(companies,file_loc){
  
  arguments <- as.list(match.call())
  floc = eval(arguments$file_loc)
  companies = eval(arguments$companies)
  i <- 1
  for(company in companies){
     path <- paste(floc[company])
     temp_name <- data.frame()
     temp_name <- read.csv(path,stringsAsFactors = FALSE)
     comp_name <- rep(companies[[i]], nrow(temp_name))
     temp_name$company <- as.character(as.factor(comp_name))
     stock <<- rbind(stock,temp_name)
     i <- i + 1  
     }
}

load_data(companies,file_loc)

```

Summary of Data set
-------------------
### What is the structure of your dataset?



### What is/are the main feature(s) of interest in your dataset?


 
```{r echo=FALSE, message=FALSE, warning=FALSE, Summary}
str(stock)
summary(stock)
```

# Data Wrangling Section

```{r}
data_wrangle <- function(){
      
      stock <<- subset(stock,stock$Close != "null")
      
      stock$Open <<- as.numeric(as.character(stock$Open))
      stock$Close <<- as.numeric(as.character(stock$Close))
      stock$High <<- as.numeric(as.character(stock$High))
      stock$Low <<- as.numeric(as.character(stock$Low))
      stock$Adj.Close <<- as.numeric(as.character(stock$Adj.Close))
      stock$Volume <<- as.numeric(stock$Volume)
      stock$company <<- stock$company
      stock$Date <<- as.Date(stock$Date)

}

data_wrangle()


str(stock$Open)
summary(stock)

```




```{r}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

# Univariate Plots Section
```{r}
# Function to calculate different financial indexes
apply_sma = function(var, n, company_df)
{
  arguments <- as.list(match.call())
  y = eval(arguments$var, company_df)
  
  sma <- SMA(y,n)
  sma
}

apply_ema = function(var, n, company_df)
{
  arguments <- as.list(match.call())
  y = eval(arguments$var, company_df)
  
  ema <- EMA(y,n)
  ema
}


apply_evwma = function(var, v, n, company_df)
{
  arguments <- as.list(match.call())
  y = eval(arguments$var, company_df)
  z = eval(arguments$v, company_df)
  
  evwma <- EVWMA(y,z,n)
  evwma
}

apply_macd = function(var, a, b, c, m, company_df)
{
  arguments <- as.list(match.call())
  y = eval(arguments$var, company_df)
  
  macd <- MACD(y,a,b,c,m)
  macd
}

apply_rsi = function(var,n,m,company_df)
{
  arguments <- as.list(match.call())
  y = eval(arguments$var, company_df)
  
  rsi <- RSI(y,n,m)
  rsi
}

apply_bbands = function(df)
{
  
  bbands <- BBands(df)
  bbands
}

apply_stoch = function(df)
{
  stoch <- stoch(df)
  stoch
}

apply_smi = function(df)
{
  smi <- SMI(df)
  smi
}

apply_wpr = function(df)
{
  wpr <- WPR(df)
  wpr
}

apply_aroon = function(var, n, company_df)
{
  arguments <- as.list(match.call())
  y = eval(arguments$var, company_df)
  
  aroon <- aroon(y,n)
  aroon
}



hcl_df <- subset(stock,(stock$company=='hcl'))
hcl_df$SMA <- apply_sma(Close, 30, hcl_df)
hcl_df$EMA <- apply_ema(Close, 30, hcl_df)
hcl_df$EVWMA <- apply_evwma(Close, Volume, 30, hcl_df)
hcl_df$MACD <- apply_macd(Close, 12,26,9,'SMA', hcl_df)
hcl_df$RSI <- apply_rsi(Close, 14,'SMA',hcl_df)
hcl_df$BBANDS <- apply_bbands(hcl_df[,c("High","Low","Close")])
hcl_df$STOCH <- apply_stoch(hcl_df[,c("High","Low","Close")])
hcl_df$SMI <- apply_smi(hcl_df[,c("High","Low","Close")])
hcl_df$WPR <- apply_wpr(hcl_df[,c("High","Low","Close")])
hcl_df$AROON <- apply_aroon(Close, 30, hcl_df)


```





```{r}

hcl_df <- subset(stock,(stock$Company=='hcl'))

A <- ggplot(hcl_df)+
  geom_line(aes(Date,Close))+
  geom_line(aes(Date,SMA(hcl_df['Close'],30)),col="blue")+
  geom_line(aes(Date,EMA(hcl_df['Close'],30)),col="red")+
  geom_line(aes(Date,EVWMA(hcl_df['Close'],Volume, 30)),col="orange")+
  ylab("Closing Price (Rs.)")+
  scale_x_date(date_breaks = "month", date_labels = "%Y-%m")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

macd <- MACD(hcl_df['Close'],12,26,9,matype=SMA)
macd_df <- as.data.frame(macd)
macd_df$Date <- hcl_df$Date
B <- ggplot(subset(macd_df,!is.na(signal)),aes(Date,signal))+
    geom_line()+
  scale_x_date(date_breaks = "month", date_labels = "%Y-%m")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("MACD Signal")

RSI <- RSI(hcl_df['Close'], n = 14,SMA)
RSI_df <- as.data.frame(RSI)
RSI_df$Date <- hcl_df$Date
C <- ggplot(subset(RSI_df,!is.na(RSI)),aes(Date,RSI))+
    geom_line()+
  scale_x_date(date_breaks = "month", date_labels = "%Y-%m")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("RSI")

bbands <- BBands( hcl_df[,c("High","Low","Close")] )
bbands_df <- as.data.frame(bbands)
bbands_df$Date <- hcl_df$Date
D <- ggplot(subset(bbands_df,!is.na(mavg)))+
    geom_line(aes(Date,dn))+
    geom_line(aes(Date,mavg),col="red")+
    geom_line(aes(Date,up),col="blue")+
  scale_x_date(date_breaks = "month", date_labels = "%Y-%m")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Bollinger")

stochOSC <- stoch(hcl_df[,c("High","Low","Close")])
stochWPR <- WPR(hcl_df[,c("High","Low","Close")])
stochOSC_df <- as.data.frame(stochOSC)
stochWPR_df <- as.data.frame(stochWPR)
stochOSC_df$Date <- hcl_df$Date
stochWPR_df$Date <- hcl_df$Date
stochOSC_df$stochWPR <- stochWPR_df$stochWPR

ggplot(subset(tail(stochOSC_df,100)))+
    geom_line(aes(Date,fastK))+
    geom_line(aes(Date,stochWPR), col="blue")+
    geom_line(aes(Date,1-stochWPR), col="red",linetype="dashed")
#main="Fast %K and Williams %R", 
#ylab="",
#ylim=range(cbind(stochOSC, stochWPR), na.rm=TRUE) )
#lines(tail(stochWPR, 100), col="blue")
#lines(tail(1-stochWPR, 100), col="red", lty="dashed")


ar <- aroon(hcl_df['Close'],30)
ar_df <- as.data.frame(ar)
ar_df$Date <- hcl_df$Date
E <- ggplot(subset(ar_df,!is.na(oscillator)),aes(Date,oscillator))+
    geom_line()+
  scale_x_date(date_labels = "%Y-%m",
               breaks = c(as.Date("2013-06-01"),
               seq(from=as.Date("2013-03-01"), to=as.Date("2018-09-04"), by="months"),
               as.Date("2018-04-04")))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ylab("Aroon Oscillator")
               
#nf <- layout(matrix(c(1,1,2,2), 2, 2, byrow = TRUE))

multiplot(A,B,C,D,E,cols=1)


```

We are going to analyze different variables that would be influencing to make
 the loans delinquent.

```{r echo=FALSE, message=FALSE, warning=FALSE, Loan_Status}
# Checking distribution of loan status
ggplot(data=loans,
       aes(LoanStatus))+
  geom_bar(fill="orange",aes(y=(..count..)/sum(..count..)))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ylab("Proportion of Loan Status")+
  ggtitle("Status of Loans")

```

4.4% of loans are defaulted. This is the data for loans over the years 2006 to
2014. This is lower than the current average in banking where the per year
 default rate is close to 1%.
 
```{r echo=FALSE, message=FALSE, warning=FALSE, Term}
loans$Term.f <- as.factor(loans$Term)
ggplot(data=loans %>% group_by(Term.f) %>% summarise(n=n()),
       aes(x=factor(1),y=n,fill=Term.f))+
  geom_bar(width=1,stat="identity")+
  coord_polar(theta = "y")+
  theme_void()+
  ggtitle("Distribution of Terms")
```


The loans are of 3 terms - 12,36 and 60. More than 90% of the loans are of 36
 months term. 
```{r echo=FALSE, message=FALSE, warning=FALSE, Listing_Category}
lab <- c("Not Available","Debt Consolidation","Home Improvement",
         "Business","Personal Loan","Student Use","Auto","Other",
         "Baby&Adoption","Boat", "Cosmetic Procedure", "Engagement Ring", 
         "Green Loans", "Household Expenses", "Large Purchases",
         "Medical/Dental","Motorcycle", "RV", "Taxes", 
         "Vacation", "Wedding Loans")

ggplot(data=loans,
       aes(factor(ListingCategory)))+
  geom_bar()+
  scale_x_discrete(labels = lab)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  xlab("Type of Loan")+
  ggtitle("Distribution of Loan Category")


```


More than 50% of the loans taken are under Debt Consolidation and the rest are 
under personal loans. 

```{r echo=FALSE, message=FALSE, warning=FALSE, CreditRating}
ggplot(data=subset(loans,!(CreditRating  %in% c("","NC"))),
       aes(CreditRating))+
  geom_bar(fill="orange",aes(y=(..count..)/sum(..count..)))+
  ylab("Proportion of credit ratings")+
  ggtitle("Distribution of Credit Ratings")
```


There are 7 types of Credit Ratings. 'AA' is the top most rating assigned to an
individual and 'HR' is the lowest rating assigned to an individual. Proper has 
largest 'C' rating having close to 25000 borrowers. The portfolio of credit 
ratings for prosper is uniform with 'HR' rating having low population and on the
 other end of spectrum 'AA' rating is also low.
 
```{r echo=FALSE, message=FALSE, warning=FALSE, IncomeRange}
ggplot(data=subset(loans,(IncomeVerifiable=="True")),aes(IncomeRange))+
  geom_bar()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Distribution of Income Range")
```


Mid income group from 25K to 75K has the largest number of borrowers. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Occupation}
ggplot(data=subset(loans_Occ,Occupation != "" ),
       aes(x=Occupation))+
   geom_bar()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Distribution of Occupation")
```


This is top 10 occupation going for loans from Prosper. Professionals and 
business people are the occupations having most numbers.

```{r echo=FALSE, message=FALSE, warning=FALSE, TotalInquiries}
ggplot(data=subset(loans,!is.na(TotalInquiries)),
    aes(x=TotalInquiries))+
    geom_histogram(binwidth = 4)+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
    xlim(0,50)+
  ggtitle("Distribution of Total Inquiries")

BoxPlots <- subset(loans,!is.na(TotalInquiries) &
                    TotalInquiries >= 0 &
                    TotalInquiries < 60 &
                    DelinquenciesLast7Years < 25 &
                    !is.na(LenderYield) )

boxplot(BoxPlots$TotalInquiries)

```


Most of the Inquiries are between 0 and 20. More the number of Inquiries more 
probability that loan may get default. The inquiries may be due to loan 
applications or credit card application. But Inquiries beyond 20 is not a 
good sign.

```{r echo=FALSE, message=FALSE, warning=FALSE, Delinquency}
boxplot(BoxPlots$DelinquenciesLast7Years)
```


Most of the accounts are show 0 delinquency in last 7 years. But Delinquency is
not a good sign for a borrower.

```{r echo=FALSE, message=FALSE, warning=FALSE, LenderYield}
boxplot(BoxPlots$LenderYield)
ggplot(data=subset(loans,!is.na(LenderYield)),
       aes(x=LoanGrade,y=LenderYield))+
   geom_boxplot()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Loan Grade/Lender Yield")
```


Lender Yield on bad loans is more than good loans.


### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
Gender of the borrower could have been helpful; Company the borrower is 
working for.

### Did you create any new variables from existing variables in the dataset?
- Created a LoanGrade Variable that was marked as 'Good' for current and paid 
off loans and 'Bad' for otherwise.
- Another data frame was created based on top 10 occupation taking loans.
- LoanOriginationQuater field is separated to Year and Quarter
- Created new variable for collected interest minus the fees

### Of the features you investigated, were there any unusual distributions? \
- Monthly income showed very large values

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, LoanGrade}
ggplot(data=subset(loans,IsBorrowerHomeowner!=""),
       aes(x=IsBorrowerHomeowner,fill=LoanGrade))+
  geom_bar(position="dodge")+
  ggtitle("Borrower is Homeowner for Good or Bad Loans")
```


As per my understanding, people who are home owners might bemaking more default
on loans. This is because it might be due to home loans on them too. So 
starting with checking if Borrower defaults more when he is home owner.
Borrower who are not homeowners seems to be bit more on default count. 

```{r echo=FALSE, message=FALSE, warning=FALSE, MonthlyIncome}
ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 60000)),
       aes(LoanGrade,StatedMonthlyIncome))+
  geom_boxplot()+
  ggtitle("Box plot - Income/Loan Grade")+
  ylab("Monthly Income(USD)")

ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 60000 &
                            !(CreditRating  %in% c("","NC")))),
       aes(CreditRating,StatedMonthlyIncome))+
  geom_boxplot()+
  ggtitle("Box plot - Rating/Income")+
  ylab("Monthly Income(USD)")

ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 100000)),
       aes(LoanOriginationYear,StatedMonthlyIncome))+
  geom_boxplot()+
  ggtitle("Box plot - Year/Income")+
  ylab("Monthly Income(USD)")

```

Next important point was to check the individual's salary monthly and its
change over time or people having different credit.As expected, people having
better salary and better credit tend to default less.

Borrowers ending up with good loans have slightly better monthly income. 
As expected lower rating have lower income and as the rating increases,
the income increases too.
Over the years portfolio is between 3000 and 5500 for 50% of
the borrowers. 

```{r echo=FALSE, message=FALSE, warning=FALSE, TotalInquiries_Bi}
ggplot(data=subset(loans,!is.na(TotalInquiries) &
                     TotalInquiries < 30),
    aes(x=LoanOriginationYear,y=TotalInquiries))+
    geom_bar(stat="identity")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Total Inquiry Over Years")

ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 100000) &
                         !is.na(TotalInquiries) & 
                        TotalInquiries < 30 &
                        !(CreditRating  %in% c("","NC"))),
       aes(CreditRating,TotalInquiries,col=CreditRating))+
  geom_boxplot()+
  scale_color_brewer(palette = "RdYlGn")+
  ggtitle("Credit Rating/Inquiries")
```

Total Inquiries is an important parameter that has to be study further. Total
 Inquiries lets us know if there were inquries on borrower's financial
 condition. Hence this an important parameter to include.
Total Inquiuries for lower ratings is more than higher ratings 
which indicates that people with bad credit history have more inquries on their 
financial status.

```{r echo=FALSE, message=FALSE, warning=FALSE, BorrowerAPR}
ggplot(data=subset(loans,!is.na(BorrowerAPR)), 
    aes(x=LoanGrade,y=BorrowerAPR))+
    geom_boxplot()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Box plot - Loan Grade/APr")

ggplot(data=subset(loans,!is.na(BorrowerAPR)), 
    aes(x=LoanOriginationYear,y=BorrowerAPR))+
    geom_boxplot()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Box plot - APR over years")

ggplot(data=subset(loans,!is.na(BorrowerAPR) &
                     !(CreditRating  %in% c("","NC"))), 
    aes(x=CreditRating,y=BorrowerAPR,col=CreditRating))+
    geom_boxplot()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
   scale_color_brewer(palette = "RdYlGn")+
  ggtitle("Box plot - APR for different rating groups")
```

Borrower APR is cost of credit which is more for loans given on bad credit. This
 is a important parameter to understand the delinquency.

Borrower APR which is actually cost of credit is more for bad loans. Borrower 
APR is interest rate + other costs. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Delinquency_bi}
ggplot(data=subset(loans,!is.na(DelinquenciesLast7Years) &
                     DelinquenciesLast7Years < 15), 
    aes(x=LoanOriginationYear,y=DelinquenciesLast7Years))+
    geom_boxplot()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Box plot - Delinquencies over years")
```


Delinquency from last 7 years indicates the financial health.
Delinquency from last 7 years are mostly 0 and over the years portfolio has less
 delinquent profiles
 
 
# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, MutliVariate_1}
ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 60000 &
                            !(CreditRating  %in% c("","NC")) &
                            !is.na(CreditScore.mean)) &
                     CreditScore.mean > 500),
       aes(x=CreditRating,y=CreditScore.mean,
           colour=CreditRating))+
  geom_boxplot()+
  scale_colour_brewer(palette="RdYlGn")+
  facet_grid(LoanGrade ~ .)+
  ggtitle("Credit Score/Credit Rating for Good/Bad Loans")

ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 60000 &
                            !(CreditRating  %in% c("","NC")) &
                            !is.na(CreditScore.mean)) &
                     CreditScore.mean > 500 &
                     !is.na(InquiriesLast6Months)),
       aes(x=StatedMonthlyIncome,y=BorrowerAPR,
           colour=CreditRating))+
  geom_point()+
  facet_grid(LoanGrade ~ .)+
  geom_jitter()+
  scale_colour_brewer(palette="RdYlGn")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Total Inquiries and Credit Score 
  over the years for different loan grades \n")+
  xlab("Monthly Income (USD)")+
  ylab("Borrower APR(%age)")

```


With few exceptions, most of the high credit borrowers have excellent credit 
score. If we facet it on Good/Bad loans credit score reduces for each Credit
 Ratings when we go from good to bad.

```{r echo=FALSE, message=FALSE, warning=FALSE, MutliVariate_2}



ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 60000 &
                            !(CreditRating  %in% c("","NC")) &
                            !is.na(CreditScore.mean))&
                     CreditScore.mean > 500),
       aes(x=StatedMonthlyIncome,y=CreditScore.mean,
           colour=CreditRating))+
  geom_point(alpha=0.9)+
  scale_colour_brewer(palette="RdYlGn")+
  facet_grid(. ~ LoanGrade)+
  ggtitle("Credit Score and Income for different rating groups")



```

 Faceting again on second dimension to IsBorrowerOwner there are changes to plot
 Plot on 3rd quadrant shows that higher the income and score and being
 home owner ensures good loans.

```{r echo=FALSE, message=FALSE, warning=FALSE, MutliVariate_3}
ggplot(data=loans_selected,
       aes(y=TotalInquiries,x=StatedMonthlyIncome,
           col=CreditRating,
           size=DelinquenciesLast7Years))+
   geom_point(alpha=1/2,
              position='jitter')+
   scale_colour_brewer(palette = "RdYlGn")+
   facet_grid(LoanGrade~LoanOriginationYear)+
   theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Total Inquiries and Monthly Income
over the years of different credit ratings")

```

With years, number of loan accounts having delinquency have reduced.

```{r echo=FALSE, message=FALSE, warning=FALSE, KNN}
loans_selected.knn <- select(loans_selected, LoanGrade,
                             CreditRating, 
                             Term,
                             CreditScore.mean,
                             StatedMonthlyIncome,
                             DelinquenciesLast7Years,
                             TotalInquiries,
                             IsBorrowerHomeowner,
                             LoanOriginationYear)


loans_selected.knn$CreditRating <-
  as.numeric(factor(loans_selected.knn$CreditRating,levels = 
                      c('HR','E','D','C','B','A','AA')))

loans_selected.knn$IsBorrowerHomeowner <-
  as.numeric(factor(loans_selected.knn$IsBorrowerHomeowner
                    ,levels = 
                      c('False','True')))
loans_selected.knn$LoanOriginationYear <- 
  as.integer(loans_selected.knn$LoanOriginationYear)

loans.train <- subset(loans_selected.knn,LoanOriginationYear < 2013)
loans.test <- subset(loans_selected.knn,LoanOriginationYear >= 2013)

knn.1 <-  knn(loans.train[,-1],loans.test[,-1],loans.train[,1], k=1)

table(knn.1, loans.test[,1])
mean(knn.1==loans.test[,1])
```
# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Interesting relationships were found with following features:
1. CreditRating
2. Term
3. CreditScore.mean
4. StatedMonthlyIncome
5. DelinquenciesLast7Years
6. TotalInquiries
7. IsBorrowerHomeowner
8. LoanOriginationYear

Some very obvious relationships were found. For instance,
Short term loans were less delinquent; Better credit ratings resulted in less
defaults


### Were there any interesting or surprising interactions between features?
Interesting insights was that before 2010 even high credit ratings
resulted in defaults. Also borrowers who are not home owners have more
bad loans. I was expecting homeowners to have more. Then I had observed that
Credit Rating are in some cases not congruent to Credit Score. Even Low
credit rating have high scores more than 700.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.
Yes created a knn model with k=1. It has accuracy of 93%.

# Final Plots and Summary


### Plot One
```{r echo=FALSE, Plot_One}

ggplot(data=subset(loans,!is.na(BorrowerAPR) &
                      !(CreditRating %in% c("","NC"))), 
    aes(x=LoanOriginationYear,y=BorrowerAPR,col=CreditRating))+
    geom_point(alpha=1/2)+
    geom_jitter()+
    scale_colour_brewer(palette="RdYlGn")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Loan Grade/APR")


```

### Description One
What we are witnessing is scatter plot between APR and Origination Year. It is 
evident form the plot that borrower APR increases as the rating goes down. Also,
the plot shows that borrower APR was disparate during initial years of Propser
business for each rating. For example, for rating 'B' the APR was between
10% and 25% in year 2008. In 2014 the distribution has become more uniform.
The reason I am exploring this is because this can be one of the reasons for 
loans getting default. Incorrect distribution of interest rate across 
different rating. 

In the next plot we are going to see other parameters which would influence the
loan default. In the above plots we seen some of them:
- Credit Score
- Inquiries
- Monthly Income
- Employement Status




### Plot Two
```{r echo=FALSE, Plot_Two}

ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 60000 &
                            !(CreditRating  %in% c("","NC")) &
                            !is.na(CreditScore.mean)) &
                     CreditScore.mean > 500 &
                     !is.na(InquiriesLast6Months) & 
                     !is.na(EmploymentStatus)),
       aes(x=StatedMonthlyIncome,y=BorrowerAPR,
           colour=CreditRating))+
  geom_point(alpha=1/4)+
  facet_grid(. ~ LoanGrade)+
  geom_jitter()+
  scale_colour_brewer(palette="RdYlGn")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Total Inquiries and Credit Score 
over the years for different loan grades \n")+
  xlab("Monthly Income (USD)")+
  ylab("Borrower APR(%age)")


```

The bad loans have the APR randomly distributed. However, with good loans the 
APR is properly structured and has less variance. There are points where APR is
low and still loan has become delinquent.
Now, the question is: How to  know the circumstances in which the loan has 
become default.I have tuned the parameters as:
 Inquiries = 0
 Credit Score > 800
 
 for the next plot. The next plot has to be compared with the above plot. Idea 
 is to minimize the points in Bad category to come to a conclusion that these 
 are the parameters which can be tuned to see that loans dont default.

 
```{r}
ggplot(data=subset(loans,(IncomeVerifiable=="True" & 
                            StatedMonthlyIncome < 60000 &
                            !is.na(CreditScore.mean)) &
                     CreditScore.mean > 750 &
                     !is.na(InquiriesLast6Months) &
                     InquiriesLast6Months == 0 ),
       aes(x=StatedMonthlyIncome,y=BorrowerAPR,
           colour=CreditRating))+
  geom_point()+
  facet_grid(LoanGrade ~ .)+
  scale_colour_brewer(palette="RdYlGn")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Monthly Income and APR for different loan grades
          in case of high rate loans('A','AA')\n")+
  xlab("Monthly Income (USD)")+
  ylab("Borrower APR (%age)")


```

The bad loans have considerably decreased by setting the parameters as mentioned
above. So the conclusion is that in order to reduce the probablity of loan 
default we can
- Either borrower with rating 'AA','A'
- OR for lower rating like 'B','C','D', credit score to be more than 750 and no.
 of enquiries = 0

 
# Reflection
Prosper loans is peer to peer lending company that provides loans for debt
consolidation and personal loans. We have data of Prosper loans starting from
2005 until 2014. Our objective is to find an answer to very important question
that whether particular loan will be delinquent or not. There were several
parameters related to nature of loan and broower profile. 

We saw that a borrower with good ratings and score is less tend to default but
lender yield is low. It is seen that high income and good rating of a borrower
makes him a safe choice for proper. From the data the proportion of good vs bad
loans is almost same. So it does not provides a good indicator to say that a 
particular occupation more prone to default. If an individual is prone to 
delinquencies in the past there are indications that the same can happen in 
future.

Loans having large term tend to default more. 

